# Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
```

All of our data are based on the 2016 Medicare & Medicaid EHR Incentive Program since the adoption of EHR is an ongoing basis: a provider who adopted and used EHR during the 2014 program, for example, is likely to still be in the program the following years. So theorethically, the 2016 program would be the most cumulative source of providers adopting EHR. As mentioned in [Related Work](https://euniceyeh.github.io/EHR-Project/work.html), the adoption of EHR has been consistently rising over the years across the U.S., so there is also no interest in looking at the use over time when we can just look at the latest collection of data. And since 2017 isn't quite over yet, it wouldn't be fair to start looking at the usage of EHR within the 2017 program, so the most recent completed data would be on the 2016 program.

## Eligible Professionals (EPs)

To answer the first of our [Initial Questions](https://euniceyeh.github.io/EHR-Project/questions.html) as to whether or not an EP's age, education, practice location, etc. may be related to their use of EHR in the incentive program, we must be able to find demographics data at the practitioner level.

1. Data Source & Metadata

And here we are: readily available on [Data.Medicare.gov](https://data.medicare.gov/Physician-Compare/Physician-Compare-National-Downloadable-File/mj5m-pzi6) are the demographics of professionals eligible for the Medicare & Medicaid EHR incentive program at the practitioner level.

```{r load-EPs-data, cache=TRUE}
# EPs <- read.cvs("https://data.medicare.gov/api/views/mj5m-pzi6/rows.csv?accessType=DOWNLOAD")
EPs <- read.cvs("https://data.medicare.gov/api/views/mj5m-pzi6/rows.csv", nrows = 10)
```

2. Exploring & Understanding the Data

```{r}
# str(EPs)
```


4. Cleaning Decisions

  - mention separating out the practitioners who are affiliated with any hospital because then the choice of EHR use is no longer at the practitioner-level rather at the hospital-level. Their demographics would no longer be informative as to whether or not they use EHR, it would then have to be the demographics of the hospitals that could inform the use of EHR at the hospital-level. Thus, we were able to find hospital demographics such as ___ on hospitals across the U.S. in general, which will be discussed in the next section within this chapter, but there's no indication of participation in the incentive program nor EHR use, so we will have to merge that hospital demographic data with the practitioners in this dataset who are affiliated with any one hospital. (note that only first set of many `hosp_afl_` columns is relevant).
  
```{r save-EPs-data}
# separate out EPs to those not affiliated with any hospital = EPs --> will be used for 
# EPs_cleaned <- data.frame(x = 1, y = 1:10)
# saveRDS(EPs_cleaned, file = "./data/EPs.rds")

# remember don't need to save the hosp_afl subset of the data because it will be merged soon within the chapter (will save the merged version for analysis)
```
  
  
5. Purpose (why/how is this data useful?)

  - `EPs` data will be used to analyze the potential effects of practitioner demographics on the use of EHR
  - `hosp_afl` is to be merged with the hospital demographics data discussed in the next section
  - both are to be merged with the EHR vendor and product information to explore these demographics by specific types of EHR used.



## Eligible Hospitals (Hosp)

General hospital demographics such as the number of staffed beds, average *(?)* number of discharges, average *(?)* number of days patients stay, and gross patient revenue are all publicly accessible on the [American Hospital Directory](https://www.ahd.com/states/) organized by U.S. states. Since there is not a `.csv` file ready for us to download and the hospital demographics are already neatly summarized in a table for each state, we decided that we could do some web scraping for this.

Since we have separated 

1. Data Source & Metadata

The link to each U.S. state's summary table of hospital demographics by hospital name in that state begins with the same URL: `https://www.ahd.com/states/hospital_` and ends with the state's abbreviation plus a `.html`. For example, a table of hospital demographics for each hospital in Massachusetts can be directly scraped from the link, [https://www.ahd.com/states/hospital_MA.html](https://www.ahd.com/states/hospital_MA.html).

2. Exploring & Understanding the Data

  - looks like there are some hospitals with really low values - turns out they indicate missing data with zeroes.

3. Cleaning Decisions

  - convert all zeroes to missing so that the zero values do not misinform summary statistics, and we highly doubt that the zeros are actually meaningfully zeros...*right?*
  - merge in with `hosp_afl`, which is subset of `EPs` dataset who are affiliated with a hospital, on the hospital name and city (since different hospitals in different locations can have the same name).

4. Purpose (why/how is this data useful?)

  - to analyze the potential effects of hospital demographics on the use of EHR
  - and to subsequently merge with the EHR vendor and product information to explore these demographics by specific types of EHR used.



## Vendors & Products (EHR)

Since so many professionals and hospitals across the U.S. have already adopted EHR, we realized that simplying looking at the binary outcome of EHR use would be too boring of a project. Thus, we were curious to further explore the demographics of those who do use EHR by the specific types of EHR vendor or product. Fortunately, we were able to find data to support this additional secondary/exploratory analysis.

1. Data Source & Metadata

The [Health IT Dashboard](https://dashboard.healthit.gov/datadashboard/documentation/ehr-products-mu-attestation-data-documentation.php) provides certified health IT product data from the ONC Certified Health IT Product List (CHPL) such as the unique vendors, products, and product types of each certified health IT product used as part of the Medicare EHR Incentive Program. We downloaded only the 2016 dataset, which also includes unique provider identifiers (NPI), in order to match the `EPs` dataset discussed in the first section of this chapter. As the [metadata](https://dashboard.healthit.gov/api/open-api.php?source=onc-data-documentation.csv&data=ehr-products-mu-attestation-data-documentation) explains, a provider in this dataset can be either an eligible professional (EP) and eligible hospital (Hospital), as distinguished by the `Provider_Type` column. Thus, only the `Provider_Type == 'EP'` records is merged with the subset of `EPs` who are not affiliated with any hospital, and the `Provider_Type == 'Hospital'` records is merged with the subset of `EPs` who are affiliated with a hospital, and therefore directly merged with the combined and cleaned version of the `hosp` dataset discussed in the previous section of this chapter.

```{r load-EHR-data, cache=TRUE}
EHR <- read.csv("https://dashboard.healthit.gov/datadashboard/data/MU_REPORT_2016.csv")
```


2. Exploring & Understanding the Data

  - specific product types are too dirty and not meaningful, will focus on vendor types.
  
3. Cleaning Decisions

  - merge with the subset of EPs who are not affiliated with any hospital, basically just the cleaned `EPs`
  - merge with the cleaned hosp data

4. Purpose (why/how is this data useful?)

Although the specific product types are too granular and not meaningful enough to clean, we believe that exploring the professional and hospital demographics across vendor types may still be interesting, especially if we focus on the top 10 most popular vendors or classify the smaller local vendors as one group.
